---
title: "Data Wrangling and Data Management for Public Health, Health Services and Primary Care"
output: html_notebook
---

Units of Assessment Studied: Public Health, Health Services and Primary Care - UOA 2

Data Source:
1) Results & Submissions Data
Downloaded from http://results.ref.ac.uk/DownloadSubmissions/ByUoa/02 and http://results.ref.ac.uk/DownloadResults

2) Number of Academic Staff Per University from HESA (Higher Education Statistical Agency) 2013/2014 (Not Used)
https://www.hesa.ac.uk/data-and-analysis/key-tables?year=3&by=579

Notes in Data Quality (Error notice):
https://www.hesa.ac.uk/support/data-intelligence?data_streams=5&year=3&field=All

3) REF 2014 Staff Contextual Data Provided by HESA (Approximated Figures)
https://www.hesa.ac.uk/news/18-12-2014/research-excellence-framework-data

Steps Undertaken:
Step 1 - Load Required Data Source (raw data come in Excel - save required sheet as .csv prior to loading)

Required Data:-
a) Staff summary  - UKPRN, Total Category A FTE, Total Head Count (Optional)
b) Submission Profile - UKPRN, Profile - Outputs (3 and 4 star)
c) Output - UKPRN, University Name, Output Count (calculated)
d) Scaled FTE of Eligible Staff 2013/14 - UKPRN, UOA, Scaled FTE of Eligible Staff

```{r}
# Load data for item a - Staff Summary
library(readr)
StaffSummary <- read_csv("~/03 COM6911 ITP/Data/Selected/2 - Public Health, Health Services and Primary Care (1)/PubHealth_StaffSummary.csv", 
    col_types = cols(MultipleSubmission = col_skip(), 
        TotalCategoryCHeadCount = col_skip(), 
        UnitOfAssessment = col_skip()))
View(StaffSummary)

```

```{r}
# Load data for item b - Submission Profile
library(readr)
SubmissionProfile <- read_csv("~/03 COM6911 ITP/Data/Selected/2 - Public Health, Health Services and Primary Care (1)/PubHealth_SubmissionProfile.csv", 
    col_types = cols(MultipleSubmission = col_skip(), 
        UnitOfAssessment = col_skip()))
View(SubmissionProfile)
```

```{r}
# Load data for item c - Output
library(readr)
Output <- read_csv("~/03 COM6911 ITP/Data/Selected/2 - Public Health, Health Services and Primary Care (1)/PubHealth_Output.csv", 
    col_types = cols(ArticleNumber = col_skip(), 
        CitationsApplicable = col_skip(), 
        CrossReferToUoa = col_skip(), DOI = col_skip(), 
        DoubleWeightingStatement = col_skip(), 
        EnglishAbstract = col_skip(), FirstPage = col_skip(), 
        ISBN = col_skip(), ISSN = col_skip(), 
        Identifier = col_skip(), IsInterdisciplinary = col_skip(), 
        IsNonEnglishOutput = col_skip(), 
        IsReserveOutput = col_skip(), Issue = col_skip(), 
        MultipleSubmission = col_skip(), 
        PatentNumber = col_skip(), Place = col_skip(), 
        ProposeDoubleWeighting = col_skip(), 
        Publisher = col_skip(), ResearchGroup = col_skip(), 
        URL = col_skip(), UnitOfAssessment = col_skip(), 
        Volume = col_skip()))
View(Output)
```

```{r}
# Load data for item d - Scaled FTE for Eligible Staff for REf 2013/14

library(readr)
Eligible_FTE_Staff <- read_csv("~/03 COM6911 ITP/Data/Selected/Eligible_FTE_Staff.csv", 
    col_types = cols(`HE provider` = col_skip(), 
        INSTID = col_skip(), `Multiple \nsubmission letter` = col_skip(), 
        `Region \nof provider` = col_skip(), 
        X9 = col_skip()))
View(Eligible_FTE_Staff)
```

Comment - Data loading error - Need to be pre-processed for item d above. Based on viewed data, need to delete 1st row and last 4th row of the table, but these error will be omitted as the data will be pre-processed to its own UOA only. 

++++++++

Step 2 - Pre-processed Data

a) Filter 'output' for score ranking 4 star and 3 star for "SubmissionProfile", and filter respective UOA for "Eligible_FTE_Staff"

```{r}
# Function to filter data frame as per required value

filter_table <- function (input_table, input_column, filter_by) {
  # input_table - Name of data frame to be filtered
  # input_column - Name of column for which criteria to be filtered
  # filter_by - value of filtered data - use format 'VALUE'
  # output_name - Name of output variable after filtering pre-processing
  subset.data.frame(input_table, subset = input_column == filter_by)
}

# Filter submission profile for 'Outputs' value ONLY
Output_Rank <- filter_table(SubmissionProfile, SubmissionProfile$Profile, 'Outputs')
```

```{r}
#Filter "Eligible_FTE_Staff for its UOA ONLY"

Scaled_FTE <- filter_table(Eligible_FTE_Staff, Eligible_FTE_Staff$`Unit of 
assessment number`, '02')

```

b) Calculate how many papers submitted by each university & how many categories of output data type submitted for each UOAs

```{r}
# Function to calculate count for categorical variable

# Need to install plyr packages first and read the library
#install.packages("plyr")
library(plyr)

calculate <- function (input_table, var_name){
  # var_name format -> "NAME" 
  out <- as.data.frame(count(input_table, c(var_name)))
}

# Calculate how many papers submitted by each university
Submissions_per_univ <- calculate(Output, "UKPRN")

# Calculate how many categories of output type submitted for each UOAs
OutputType_per_uoa <- calculate(Output, "OutputType")

```

c) Sum & find average of 3 star and 4 star ranking 

```{r}
# Function to sum and find average 3 star and 4 star of chosen UOAs 

sum_average <- function (input_data1, input_data2, n = 2){
  sum = input_data1 + input_data2 
  avg =  sum / n
}

Output_Rank$avg3_4star <- sum_average(Output_Rank$FourStar, Output_Rank$ThreeStar)

```

d) Calculate % (submitted FTE (Category A) vs eligible FTE) per university 

```{r}
# Function to calculate % (submitted FTE vs eligible FTE)

percent_FTE <- function (submitted_FTE, eligible_FTE){
  percent = (submitted_FTE / eligible_FTE) * 100
}

# Merge submited FTE with eligible FTE due to dissimilar row size in both (103 vs 89)
FTE <- merge.data.frame(StaffSummary, Scaled_FTE)

FTE$Percent_Submitted <- percent_FTE(FTE$TotalCategoryAFte, FTE$`Scaled FTE 
of eligible staff`)

```

